---
title: "Kaggle: Spaceship Titanic"
execute: 
  warning: false
  message: false
---

## Preliminaries

```{r}
library(tidyverse)
library(caret)
library(rpart)
library(rpart.plot)
library(visdat)
set.seed(0)
```

```{r}
main.dir <- "C:\\Users\\erikv\\Downloads\\spaceship-titanic\\"
train.csv <- paste0(main.dir,"train.csv")
test.csv <- paste0(main.dir,"test.csv")

train <- read_csv(train.csv)
test <- read_csv(test.csv)
```

## Begin Feature Engineering

From the data descriptions the `PassengerId` variable also indicates the group a passenger was traveling with. Similarly, the `Cabin` variable indicates the deck, room number, and side of the ship a passenger was on.

```{r}
# add in group variable based on PassengerId
train$Group <- substr(train$PassengerId, 1, 4)

# separate Cabin variable into deck, number, side from "/" delimiter
train <- train %>% 
  separate_wider_delim(Cabin, delim = "/", names = c("Deck", "Number", "Side"))

train <- train %>% 
  mutate_if(is.character, as.factor)
```

Next, we need to check the types of each columns to determine if any changes need to be made.

```{r}
summary(train)
```

Let's get a split from the train set to test model accuracy on.

```{r}
inds <- sample(nrow(train), .7*nrow(train))

train.set <- train[inds,]
test.set <- train[-inds,]
```

```{r}
space_dt1 <- rpart(
  formula = Transported ~ HomePlanet + CryoSleep + Deck + Number + 
    Side + Destination + Age + VIP + RoomService + FoodCourt + 
    ShoppingMall + Spa + VRDeck + Group,
  data    = train.set,
  method  = "class",
  cp = .008,
  na.action = na.pass
)
```

```{r}
space_dt1
```

```{r}
rpart.plot(space_dt1)
```



```{r}
pred.tree = predict(space_dt1, test.set, type = "class")
```

```{r}
table(pred.tree,test.set$Transported)
```

