---
title: "Week of 1/9/2024"
execute: 
  warning: false
  message: false
---

## Preliminaries

First we load up the datasets for this week. 

```{r}
library(tidyverse)
library(viridis)
library(lubridate)

# tuesdata <- tidytuesdayR::tt_load('2024-01-09')

```

In the discussion of the dataset the first chapter of (https://www.amazon.com/Outliers-Story-Success-Malcolm-Gladwell/dp/0316017930)[Malcolm Gladwell's Outliers] is brought up and it's mentioned that in Canadian Junior Hockey there is a higher likelihood for players to be born in the first quarter of the year. We are given data on NHL players and their teams as well as data on births in Canada from 1991 to 2002. Before diving into the NHL data, let's look at how births are distributed by quarter year over year.

```{r}
# canada_births_1991_2022 <- tuesdata$canada_births_1991_2022
# nhl_player_births <- tuesdata$nhl_player_births
# nhl_rosters <- tuesdata$nhl_rosters
# nhl_teams <- tuesdata$nhl_teams
setwd('data')
canada_births_1991_2022 <- read_csv('canada_births_1991_2022.csv')
nhl_player_births <- read_csv('nhl_player_births.csv')
nhl_rosters <- read_csv('nhl_rosters.csv')
nhl_teams <- read_csv('nhl_teams.csv')
```

